\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\input{00-imports}

\input{00-defs}

\input{00-settings}

% ### uncomment to produce preprint IEEE copyright notice ###
%\AtBeginShipout{%
%  \AtBeginShipoutUpperLeft{%
%    \put(\dimexpr.5\paperwidth-9cm\relax,-1cm){%
%      \parbox{\textwidth}{\centering\footnotesize%
%        © 2023 IEEE. Personal use of this material is permitted. Permission from IEEE must be obtained for all other uses, in any current or future media, including reprinting/republishing this material for advertising or promotional purposes, creating new collective works, for resale or redistribution to servers or lists, or reuse of any copyrighted component of this work in other works.}%
%    }%
%  }%
%}

\begin{document}
	
\title{\LARGE \bf
LRHControl: A Software Ecosystem for Research in Reinforcement Leaning-based Receding Horizon Control\authorrefmark{3}
}

\author{
  \authorblockN{%
    Andrea Patrizi\authorrefmark{1}\authorrefmark{2}, Carlo Rizzardo\authorrefmark{1}, 
    and Nikos G. Tsagarakis\authorrefmark{1}
  }
}

\maketitle

\begingroup\renewcommand\thefootnote{\authorrefmark{2}}
	\footnotetext{Department of Informatics, Bioengineering, Robotics and Systems Engineering, Università di Genova, Via All'Opera Pia 13, 16145 Genova.}
\endgroup

\begingroup\renewcommand\thefootnote{\authorrefmark{1}}
	\footnotetext{Humanoids and Human-Centred Mechatronics (HHCM), Istituto Italiano di Tecnologia (IIT), Via San Quirico 19d, 16163 Genova.}
\endgroup

\begingroup\renewcommand\thefootnote{\authorrefmark{3}}
\footnotetext{This project has received funding from the European
	Union’s Horizon Europe Framework Programme under grant
	agreement No 101070596
}
\endgroup

\setlength{\textfloatsep}{12.0pt plus 8.0pt minus .0pt}

\begin{abstract}
Robotics research in locomotion is undergoing a transformative shift towards the use of learning-based tools. Learning methodologies have been shown to be capable of remarkable robustness and performance even when applied to real-world environments; however, they present limitations in interpretability, safety guarantees and sample efficiency. For this reason, it is the authors' belief that more classical control approaches should not be disregarded yet. We thus advocate for a hybrid approach, combining offline data-based policy design through Reinforcement Learning (RL), with classical online Motion Planning, via Receding Horizon Control (RHC). Even though this kind of hybrid approaches are not entirely new, to the authors' knowledge, there is no specific tool currently available for research in this domain. To this purpose, we developed a modular software ecosystem, hereby briefly presented in its main components and features. 
%We care to stress that the framework is currently under active development, and features might not be stable or could be lacking. 
To facilitate its usability and diffusion, we made all the core components open source under the GPLv2 license. Furthermore, to showcase the potential of our framework and approach, we briefly present a proof-of-concept example combining a high-level RL agent coupled with a lower-level MPC controller for the execution of a simple locomotion task on a simulated quadruped robot.
\end{abstract}

\IEEEpeerreviewmaketitle

\input{01-overview}
%\input{02-rlrhc_formulation}
%\input{03-rlrhc-shortcomings}
\input{04-lrhc_formulation}
\input{05-implementation}
\input{06-examples}
\input{07-future_work}
\bibliographystyle{ieeetr}
\bibliography{bibliography/refs}

\end{document}
